---
title: "Introduction to *R* and *Markdown*"
author: "Mariusz Rafało"
output:
  html_document: default
  html_notebook: default
editor_options: 
  chunk_output_type: console
---

*Support material for class: Data Visualisation and Presentation at Warsaw School of Economics*

&copy; *Copyright Mariusz Rafało*

---

####Summary
This document presents basic concepts of data exploration in *R*.

####Data
All data used in this document come from *R* package **MASS**.


####Introduction  

Environment setup

```{r setup, include=FALSE, ECHO=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, echo=TRUE, warning=FALSE, message=FALSE)
```

Libraries used in document

```{r}
library(MASS)
```

####Examples

**Example 1:** Cars sales in USA, in 1993 - data sample

```{r}

head(Cars93[,c(1:4)], n=5)
```


Sample plots

```{r}

plot(Cars93[,c("Price","Passengers","EngineSize")])

```

**Example 2:** Ship malfunctions - data sample

```{r}
head(ships, n=7)

```

Sample plots

```{r}

plot(ships$incidents)

plot(ships$incidents~ships$year, xlim=c(0,100))

plot(density(ships$incidents), lwd=4, lty=5, col=22)

```


**Example 3:**Exam results of University of Adelaide students

```{r}

data(survey)

head(survey[,c(1:4)],n=5)

table(survey$Sex,survey$Exer)

table(survey$Sex,factor(survey$Exer,levels=c("None","Some","Freq")))

round(prop.table(table(survey$Sex,factor(survey$Exer,levels=c("None","Some","Freq")))),2)


```

#### Sample data analysis:Exam results of University of Adelaide students

Data types

```{r}
str(survey)
```

Basic statistics

```{r}
summary(survey)
```

Basic plots

```{r}
plot(survey$Sex)

pie(table(survey$Sex))
```


Graphical representation of relationship between *sex* and *height* variables

```{r}
plot(survey[c("Sex","Height")])

plot(survey[c("Sex","Exer")])
```

Graphical representation of relationship between *sex* and *smoke* variables

```{r}

liczba.smoke <- table(survey$Smoke)
barplot(liczba.smoke, col=1:length(liczba.smoke))

plot(survey$Sex,survey$Smoke)

```

...continued...

```{r}
par(mfrow=c(2,1))

liczba.sex <- table(survey$Sex)
barplot(liczba.sex, col=c(12,13))

liczba.hand <- table(survey$W.Hnd)
barplot(liczba.hand, col=c(14,15))

par(mfrow=c(1,1))

```

*table* function example

```{r}

liczba.tabela <- table(survey$Smoke,survey$Sex)

liczba.tabela

plot(liczba.tabela,col=rainbow(5))
```

*factor* function example

```{r}
survey$Smoke.ordered <- factor(survey$Smoke, levels = c("Never", "Occas", "Regul", "Heavy"))

liczba.tabela <- table(survey$Smoke.ordered,survey$Sex)

liczba.tabela

plot(liczba.tabela,col=rainbow(5))
```

Adding title and legend to plot

```{r}

barplot(liczba.tabela, beside=TRUE, 
        names=names(liczba.tabela),col=rainbow(4))
title ="Palenie tytoniu a płeć"
legend(x="topright", 
      legend=rownames(liczba.tabela), fill=rainbow(4))
```


```{r}
plot(survey$Smoke.ordered~survey$Sex,col=c(20:24),xlab="Sex", ylab="Smoke")

```

#### Selected functions from *graphics* package

**par** 
```{r}
par(mfrow=c(2,2))
par(mfrow=c(2,2))
liczba.MI <- table(survey$M.I)
barplot(liczba.MI, col=c(12,13))

liczba.Fold <- table(survey$Fold)
barplot(liczba.Fold, col=c(14,15,16))

hist(survey$Pulse, xlab="Puls", main="")
plot(survey$Pulse~survey$Sex, xlab="Płeć", ylab="Puls")

par(mfrow=c(1,1))
```

**barplot** 

```{r}
liczba.W.Hnd <- table(survey$W.Hnd)
barplot(liczba.W.Hnd, col=c(12,13))
```

**pie** 

```{r}
liczba.exer <- table(survey$Exer)
pie(liczba.exer, col=c(22,23,24))
```

**hist** 

```{r}
hist(survey$Age, xlab="Wiek", main="", breaks=6, col="green")
```

**density**

```{r}
d <- density(survey$Age)
plot(d, col="red", main="")
```
