---
title: "Introduction to spatial analysis using *R*"
author: "Mariusz Rafało"
output:
  html_document: default
  html_notebook: default
editor_options: 
  chunk_output_type: console
---

*Support material for class: Data Visualisation and Presentation at Warsaw School of Economics*

&copy; *Copyright Mariusz Rafało*

---

####Summary
This document presents basic concepts of spatial analysis performed in *R*.

####Data
All data used in this document come from *R* packages **crime** and **tmap**.

####Introduction 

Environment setup

```{r setup, include=FALSE, ECHO=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, echo=TRUE, warning=FALSE, message=FALSE)
```

Clear workspace

```{r}
rm(list=ls())
```

Libraries used

```{r}
library(ggplot2)
library(ggmap) 
library(tmap)

```

####Spatial analysis usign *tmap* package

*Base on [CRAN documentation](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-nutshell.html).* 

Data

```{r}
data(World, metro, Europe, land)
hist(Europe$well_being)
```

Basic plots

```{r} 

qtm(Europe)

qtm(Europe, 
    fill="well_being", 
    text="iso_a3", 
    text.size="area", 
    #format="Europe", 
    style="gray", 
    text.root=5, 
    fill.title="Well being in Europe", 
    fill.textNA="no data")

qtm(Europe, 
    fill="pop_est", 
    text="iso_a3", 
    text.size="AREA", 
    style="grey", 
    text.root=5, 
    fill.title="Population in Europe", 
    fill.textNA="no data")

qtm(World)

data(World)

qtm(World, 
    fill = "economy", 
    format="World_wide"
#    style="col_blind"
    )

qtm(World, 
    fill="HPI", 
    fill.n=3, 
    fill.title="Happiness index", 
    style="gray",
    fill.textNA="no data")

```

Advanced plots (usoing *tm_shape*)

```{r}
tm_shape(Europe) +
  tm_polygons(c("HPI", "gdp_cap_est"), 
              style=c("pretty", "kmeans"),
              palette=list("RdYlGn", "Purples"),
              auto.palette.mapping=FALSE,
              title=c("Happiness index", "GDP per capita")) +
  tm_style_grey()

tm_shape(Europe) +
  tm_polygons("well_being", title="Well being") +
  tm_facets("part") +
  tm_style_grey()

tm_shape(Europe[Europe$continent=="Europe",]) +
  tm_fill("part", legend.show = FALSE) +
  tm_facets("name", free.coords=TRUE, drop.units=TRUE)

```

Example

```{r}
tm_shape(land, projection="eck4") +
  tm_raster("elevation", breaks=c(-Inf, 250, 500, 1000, 1500, 2000, 2500, 3000, 4000, Inf),  
            palette = terrain.colors(9), title="Elevation", auto.palette.mapping=FALSE) +
  tm_shape(World) +
  tm_borders("grey20") +
  tm_grid(projection="longlat", labels.size = .5) +
  tm_text("name", size="AREA") +
  tm_compass(position = c(.65, .15), color.light = "grey90") +
  tm_credits("Eckert IV projection", position = c(.85, 0)) +
  tm_style_classic(inner.margins=c(.04,.03, .02, .01), legend.position = c("left", "bottom"), 
                   legend.frame = TRUE, bg.color="lightblue", legend.bg.color="lightblue", 
                   earth.boundary = TRUE, space.color="grey90")
```

#####Spatial analysis using *gmap* package

Data

```{r}

data(crime)

morderstwa <- subset(crime, offense == "murder")
kradzieze <- subset(crime, offense == "theft")

ggplot(morderstwa, aes(x=number)) + geom_density()

ggplot(morderstwa, aes(x=hour, y=number)) + geom_bar(aes(fill=day), stat="identity") + ggtitle("Murders")

ggplot(crime, aes(x=hour, y=number)) + geom_bar(aes(fill=offense), stat="identity") + ggtitle("Murders")


ggplot(crime, aes(x=hour, y=number)) + geom_bar(aes(fill=offense), stat="identity") + facet_wrap(~offense)


ggplot(morderstwa, aes(x=lon, y=lat)) + geom_point(aes(color=hour), size=2) + scale_color_gradient(low = "red", high = "yellow")

tmp <- data.frame(table(crime$hour, crime$offense))

ggplot(tmp, aes(x=as.numeric(Var1), y=Freq)) + geom_line(aes(col=Var2), size=2)

```

Google maps integration

```{r}

qmap("warsaw", zoom = 14, color = "color")

pub <- geocode("warszawa, pub lolek")

qmap("Pub Lolek", maprange = TRUE, zoom=13,
   base_layer =
     ggplot(aes(x=lon, y=lat), data = pub)) +
  geom_point(color="red", size = 4)

```

Plot charts on Google maps using *qplot*

```{r}
qmplot(lon, lat, data = morderstwa, colour = "red", size = 2, darken = .1)
qmplot(lon, lat, data = kradzieze, colour = "blue", size = 2, darken = .1)

```

Plot charts on Google maps using *ggplot*

```{r}
Houston <- qmap("houston", zoom = 14, color = "bw", legend = "topleft")

Houston + geom_point(data=morderstwa, aes(x=lon, y=lat,color=hour), size=2) + scale_color_gradient(low = "red", high = "yellow")
 
  
Houston +
  geom_point(aes(x = lon, y = lat,
                 colour = "red", size = hour),
             data = kradzieze)

Houston +
  stat_density2d(
    aes(x = lon, y = lat, fill = number, alpha = 0.2), 
    #bins = 4, 
    data = kradzieze,
    geom = "polygon") + 
  theme(legend.position="none")

Houston +
  stat_density2d(
    aes(x = lon, y = lat, fill = ..level..,
        alpha = ..level..),
    size = 2, bins = 6, data = morderstwa,
    geom = "polygon")


Houston + 
  stat_density2d(aes(x=lon, y=lat,fill="green", alpha=..level..), 
                 geom="polygon", data=kradzieze) + 

  stat_density2d(aes(x=lon, y=lat,fill="blue", alpha=..level..), 
                 geom="polygon", data=morderstwa) +
  
  scale_fill_manual(values=c("green", "blue"), name="Crime",
                     labels=c("Theft", "Murder")) +
  guides(alpha=FALSE)


```
