---
title: "Logistic regression - basic concept"
author: "Mariusz Rafało"
output:
  html_document: default
  html_notebook: default
editor_options: 
  chunk_output_type: console
---

*Support material for class: Data Visualisation and Presentation at Warsaw School of Economics*

&copy; *Copyright Mariusz Rafało*

---

####Summary
This document presents basic concepts of logistic regression performed in *R*.

####Data
Dataset *telco.csv* was downloaded from [Churn in Telecom dataset](https://rickvanderzwet.nl/trac/personal/export/167/liacs/dbdm/dbdm_3/report.pdf) article.

####Introduction 

Environment setup

```{r setup, include=FALSE, ECHO=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, echo=TRUE, warning=FALSE, message=FALSE)
```

---

Load data from *telco.csv* file

```{r}
dane <- read.csv("E:/Data sets/telco/telco.csv", header=TRUE, sep=";")
```

Divide dataset into training and test datasets 

```{r}
set.seed(101) 
sample <- sample.int(nrow(dane), floor(.75*nrow(dane)), replace = F)
dane.train <- dane[sample, ]
dane.test <- dane[-sample, ]
```

####Logistic regression


The model

```{r}

fitlogit <- glm(formula = Churn ~ AccountLength + DayMins + IntlPlan + VMailPlan  + IntlCalls + CustServCalls + EveCalls + NightCalls + VMailMessage, family = "binomial", data = dane.train) 
```

Model summary

```{r}
summary(fitlogit)

reszty = residuals(fitlogit)
hist(reszty, prob=TRUE, ylim=c(0,1.8))
lines(density(reszty), col="red", lwd=2)

shapiro.test(reszty)
```

Model diagnostics: confusion matrix

```{r}
library(caret) #confusionMatrix

predict.glm <- predict(fitlogit, newdata=dane.test, type = "response") 
predict.glm1 <- predict.glm

predict.glm1[predict.glm>0.3] <-1
predict.glm1[predict.glm<=0.3] <-0


plot(density(dane.test$Churn))
lines(density(dane.test[predict.glm1==1,]$Churn))

conf.matrix.glm <- confusionMatrix(predict.glm1, as.numeric(dane.test$Churn))

conf.matrix.glm$table
```

ROC curve

```{r}
library(pROC) #roc

roc.glm <- roc(as.numeric(dane.test$Churn),predict.glm, direction="<")

plot(roc.glm,col="red", lwd=1, main="Model 1")

```
