---
title: "Linear regression - basic concept"
author: "Mariusz Rafało"
output:
  html_document: default
  html_notebook: default
editor_options: 
  chunk_output_type: console
---

*Support material for class: Data Visualisation and Presentation at Warsaw School of Economics*

&copy; *Copyright Mariusz Rafało*

---

####Summary
This document presents basic concepts of linear regression performed in *R*.

####Data
All data used in this document come from *R* package **MASS**.

####Introduction

Environment setup

```{r setup, include=FALSE, ECHO=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, echo=TRUE, warning=FALSE, message=FALSE)
```

Libraries used in document

```{r}
library(MASS)
```

---

#### Linear regression

Put *Cars93* built-in dataset into variable *dane*
```{r}
dane <- Cars93
```

Linear regression model build: function *lm*

```{r}
lm1 <- lm(Price~Horsepower, data=dane)
```

Model summary

```{r}

summary(lm1)
```

Theoretical (model) vs real values

```{r}
par(mfrow=c(1,1))
plot(dane$Price~dane$Horsepower, xlab="Horsepower", ylab="Price")
abline(lm1, col='red')
legend("topleft", legend=c("real","model"), col=c("black","red"), pch=c(1,NA), lty=c(NA,2), bty="n")

```


Model diagnostics

```{r}
par(mfrow = c(2, 2))
plot(lm1)
par(mfrow = c(1, 1))

coefficients(lm1)

plot(dane$Price~dane$Horsepower, xlab="Horsepower", ylab="Price")
abline(lm1, col='red')
segments(dane$Horsepower, dane$Price, dane$Horsepower, lm1$fitted.values,col="red")


```

Residuals analysis

```{r}
reszty = residuals(lm1)
hist(reszty, prob=TRUE, ylim=c(0,0.11))
lines(density(reszty), col="red", lwd=2)
```

Normality verification using *Shapiro-Wilk* test

```{r}
shapiro.test(reszty)

```