---
title: "Introduction to package *lattice*"
author: "Mariusz Rafało"
output:
  html_document: default
  html_notebook: default
editor_options: 
  chunk_output_type: console
---

*Support material for class: Data Visualisation and Presentation at Warsaw School of Economics*

&copy; *Copyright Mariusz Rafało*

---

####Summary
This document presents basic concepts of data exploration in *R*, using package *lattice*.

####Data
All data used in this document come from *R* package **MASS**.

####Introduction

Environment setup

```{r setup, include=FALSE, ECHO=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, echo=TRUE, warning=FALSE, message=FALSE)
```

Libraries used in document

```{r}
library(MASS)
library(lattice)
```

Put *Cars93* built-in dataset into variable *dane*
```{r}
dane <- Cars93
```

Histogram
```{r}

histogram(~Price, xlab="Cena", data=dane, col="green", main="Histogram")

histogram(~Price|Origin, xlab="Cena", data=dane, col=c(12,16))

```

Density charts

```{r}
densityplot(~Price|Origin, xlab="Cena", data=dane, groups=DriveTrain, plot.points=FALSE, auto.key = TRUE)

```

Q-Q plots

```{r}

qqmath(~Price|Origin, dane)


qqmath(~Price, dane)

qqmath(~ Price , dane, groups = DriveTrain, f.value = ppoints(100), 
       auto.key = list(columns = 3))
```

Boxplot

```{r}

bwplot(Price~Type|Origin, dane, col="blue")

bwplot(Price~Type, dane, layout=c(1,1), col="blue")

```

Bar plot

```{r}
barchart(Type~Price|Origin, dane)

```

Dot plot

```{r}

xyplot(Price~Horsepower, dane)
```

Dot plot

```{r}

dotplot(Horsepower~Price, dane, groups=Origin, pch=c(1,2), auto.key=list(columns=2))
dotplot(Price~Horsepower, data = dane, groups = Origin, type = "o",
        auto.key = list(space = "bottom", points = TRUE, lines = TRUE))

```

3d charts examples

```{r}
cloud(Price~Horsepower*Cylinders,dane,screen = list(x = 0, y=0, z=0))
cloud(Price~Horsepower*Cylinders,dane,screen = list(x = 30, y=0, z=0))
cloud(Price~Horsepower*Cylinders,dane,screen = list(x = 30, y=-5, z=0))
cloud(Price~Horsepower*Cylinders,dane,screen = list(x = 30, y=5, z=15))
```