---
title: "Decision trees - basic concept"
author: "Mariusz Rafało"
output:
  html_document: default
  html_notebook: default
editor_options: 
  chunk_output_type: console
---

*Support material for class: Data Visualisation and Presentation at Warsaw School of Economics*

&copy; *Copyright Mariusz Rafało*

---

####Summary
This document presents basic concepts of decision tree analysis performed in *R*.

####Data
Dataset *telco.csv* was downloaded from [Churn in Telecom dataset](https://rickvanderzwet.nl/trac/personal/export/167/liacs/dbdm/dbdm_3/report.pdf) article.

####Introduction 

Environment setup

```{r setup, include=FALSE, ECHO=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, echo=TRUE, warning=FALSE, message=FALSE)
```

Load data from *telco.csv* file

```{r}
dane <- read.csv("E:/Data sets/telco/telco.csv", header=TRUE, sep=";")
```

Divide dataset into training and test datasets 

```{r}
set.seed(101) 
sample <- sample.int(nrow(dane), floor(.75*nrow(dane)), replace = F)
dane.train <- dane[sample, ]
dane.test <- dane[-sample, ]
```

Libraries specific for decision tree analysis

```{r}
library(rpart)
library(rpart.plot)
```

####Decission tree

The model

```{r}
model.dt <- rpart(Churn ~ AccountLength + DayMins + IntlPlan + VMailPlan  + IntlCalls + CustServCalls + EveCalls + NightCalls + VMailMessage, data=dane.train)
```

Basic model summary

```{r}
prp(model.dt, main="", cex=1)

plotcp(model.dt)
plotcp(model.dt, cex.lab=1.1, cex.axis=1.1)

printcp(model.dt)

```

Model visualization using *rattle* package

```{r}
library(rattle)

fancyRpartPlot(model.dt, cex=0.6)

```

ROC curve

```{r}
library(pROC) #roc

predict.dt <- predict(model.dt, newdata=dane.test, type = "vector") 

roc.dt <- roc(as.numeric(dane.test$Churn),as.numeric(predict.dt), direction="<")

plot(roc.dt,col="red", lwd=1, main="Decision tree model")

```

Model diagnostics: confusion matrix

```{r}

library(caret) #confusionMatrix

predict.dt1 <- predict.dt

predict.dt1[predict.dt>0.4] <-1
predict.dt1[predict.dt<=0.4] <-0

conf.matrix.dt <- confusionMatrix(predict.dt1, as.numeric(dane.test$Churn))

conf.matrix.dt$table

tmp <- as.data.frame(conf.matrix.dt$table)
qplot(x=Prediction, y=Reference, data=tmp, fill=Freq, geom="tile")

```


