---
title: "Churn in telecommunications - exploratory data analysis"
author: "Mariusz Rafało"
output:
  html_document: default
  html_notebook: default
editor_options: 
  chunk_output_type: console
---

*Support material for class: Data Visualisation and Presentation at Warsaw School of Economics*

&copy; *Copyright Mariusz Rafało*

---

####Summary
This document presents exploratory analysis of customer churn in relationship with several behavioral variables.  

####Data
Dataset *telco.csv* was downloaded from [Churn in Telecom dataset](https://rickvanderzwet.nl/trac/personal/export/167/liacs/dbdm/dbdm_3/report.pdf) article.

####Introduction

Environment setup

```{r setup, include=FALSE, ECHO=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, echo=TRUE, warning=FALSE, message=FALSE)
```


####Churn analysis

1. Load data from *telco.csv* file

```{r}

dane <- read.csv("E:/Data Sets/telco/telco.csv", header=TRUE, sep=";")

```


2. Graphical representation of variables: DayCalls, EveCalls, NightCalls, IntlCalls

```{r}
par(mfrow=c(2,2))

hist(dane$DayCalls, xlab="DayCalls", main="DayCalls")
hist(dane$EveCalls, xlab="EveCalls", main="EveCalls")
hist(dane$NightCalls, xlab="NightCalls", main="NightCalls")
hist(dane$IntlCalls, xlab="IntlCalls", main="IntlCalls")

par(mfrow=c(1,1))
```

3.Churn rate on pie chart

```{r}
pie(table(dane$Churn), col=c(11,112))

```

4. Number of customers in different states

```{r fig.width=15, fig.height=5}
liczba.state <- table(dane$State)
barplot(liczba.state)
plot(liczba.state,type="h", lwd=7,ylim=c(30,100), col="blue")
```

5. Call center contacts for *churn* and *noChurn* customers

```{r}
plot(as.numeric(dane$CustServCalls)~factor(dane$Churn,levels=c(0,1)), xlab="Churn", ylab="Call center calls")
```

6a. Contract length vs number of daily calls

```{r}
plot(dane$AccountLength~dane$DayCalls, xlab="AccountLength", ylab="DayCalls", pch=12)
lines(lowess(dane$AccountLength~dane$DayCall), col="red", lwd=3)

```


6b. Does average call duration affects churn?

```{r}
d0 <- density(dane[dane$Churn==0,]$DayMins)
d1 <- density(dane[dane$Churn==1,]$DayMins)
plot(d0, col="green", main="")
lines(d1,col="red")

```

6c. Call center contacts and churn rate relationship

Example 1:
```{r}
plot(dane$CustServCalls,dane$DayMins, col=dane$Churn+11, pch=16, xlab="CustServCalls", ylab="DayMins")
legend(x = 6, y = 50, legend = c("noChurn","churn"), col = c(11,12), pch = 16)
```

Example 2:
```{r}
plot(lowess(dane[dane$Churn=="1",]$CustServCalls,dane[dane$Churn=="1",]$DayMins), type="l", 
     col="red", xlab="CustServCalls", ylab="DayMins")

lines(lowess(dane[dane$Churn=="0",]$CustServCalls,dane[dane$Churn=="0",]$DayMins), col="blue")
```


