---
title: "Introduction to package *graphics*"
author: "Mariusz Rafało"
output:
  html_document: default
  html_notebook: default
editor_options: 
  chunk_output_type: console
---

*Support material for class: Data Visualisation and Presentation at Warsaw School of Economics*

&copy; *Copyright Mariusz Rafało*

---

####Summary
This document presents basic concepts of data exploration in *R*, using built-in package *graphics*.

####Data
All data used in this document come from *R* package **MASS**.

####Introduction

Environment setup
```{r setup, include=FALSE, ECHO=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, echo=TRUE, warning=FALSE, message=FALSE)
```

Libraries used in document

```{r}
library(MASS)
```

#### Selected data manipulation functions

*table* and *prop.table*
```{r}

t1 <- table(Cars93$Type)

t1

prop.table(t1)

t2 <- table(Cars93$Type, Cars93$Passengers)
t2

prop.table(t2)

prop.table(table(Cars93$Type))

prop.table(table(Cars93$Type, Cars93$Passengers))

prop.table(table(Cars93$Type))

d <- Cars93

d1 <- d[,c("Manufacturer","Type","Price")]

d[d$Manufacturer=="Toyota",c("Model","Horsepower")]
```

Density plots

```{r}
x1 <- density(d[d$Type=="Compact",]$Price)
x2 <- density(d[d$Type=="Large",]$Price)

plot(x1, main="Ceny aut", col="green", xlim=c(0,50), ylim=c(0,0.1))
lines(x2, col=156)
legend("topright",legend = c("Kompakty","Du?e"), col = c("green",156), pch = c(1,1))

```

#### Data analysis using base *plot* function


```{r}

plot(Cars93$Price, xlab="Index", ylab="Cena", pch=6, col="blue")
lines(lowess(Cars93$Price))

plot(density(Cars93$Price))
plot(Cars93$Price~Cars93$MPG.city)
plot(Cars93$Price~Cars93$MPG.city, xlim=c(0,50))

plot(density(Cars93$Price), lwd=4, lty=5, col=21)

plot(Cars93$Manufacturer, col=12)

plot(Cars93$Type~Cars93$Cylinders)

plot(Cars93[,c("Price","Passengers","EngineSize")])

plot(Cars93$Price~Cars93$Type, xlab="Typ", ylab="Cena", main="Plot title")
```

**Split screen**

```{r}
par(mfrow=c(2,2))

plot(Cars93$Price, xlab="Index", ylab="Cena", pch=2)
plot(density(Cars93$Price))
plot(Cars93$Price~Cars93$MPG.city)
plot(Cars93$Price~Cars93$MPG.city, xlim=c(0,50)) 

par(mfrow=c(1,1))

```

**Two data categories on single plot**

```{r}
d0 <- density(Cars93[Cars93$Type=="Large",]$Price)
d1 <- density(Cars93[Cars93$Type=="Sporty",]$Price)


plot(d0, col="green", main="")
lines(d1, col="red")

```

**Legend**

```{r}
d0 <- density(Cars93[Cars93$Type=="Large",]$Price)
d1 <- density(Cars93[Cars93$Type=="Sporty",]$Price)
plot(d0, col="green", main="")
lines(d1,col="red")
legend('topright',legend = c("Large","Sporty"), col = c("green","red"),pch = c(3,3))
```

**Other functions**

```{r}
boxplot(Cars93$Price)

barplot(Cars93$Price)

pie(table(Cars93$Type))


```


*cut* function

```{r}
library(plyr) #cut function

pp <- cut(Cars93$Price,3, labels = c("low", "average","high"))


pp.new <- mapvalues(pp, from = c("(7.35,25.6]", "(25.6,43.7]", "(43.7,62]"), to = c("low", "average","high"))
plot(pp, main="Car price")

```

