---
title: "Basket analysis using *apriori* algorithm"
author: "Mariusz Rafało"
output:
  html_document: default
  html_notebook: default
editor_options: 
  chunk_output_type: console
---

*Support material for class: Data Visualisation and Presentation at Warsaw School of Economics*

&copy; *Copyright Mariusz Rafało*

---

####Summary
This document presents basic concepts of basket analysis in *R*, using package *arules*.

####Data
All data used in this document are from *R* package **datasets**

####Introduction 

Environment setup

```{r setup, include=FALSE, ECHO=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, echo=TRUE, warning=FALSE, message=FALSE)
```

Libraries used in document

```{r}
library(arules)
library(arulesViz)
library(datasets)
library(igraph)
```
 
Dataset: first glance 

```{r}
data(Groceries)

inspect(Groceries[1:5])

itemFrequencyPlot(Groceries,topN=10,type="absolute")

itemFrequencyPlot(Groceries,topN=10,type="relative")
```

####Basket analysis

```{r}
rules <- apriori(Groceries, parameter = list(supp = 0.001, conf = 0.8))

top10rules <- head(sort(rules, by="lift"), 10)

options(digits=2)
inspect(top10rules)

rules_subset <- subset(rules, (rhs %in% "whole milk"))

top20rules.milk <- head(sort(rules_subset, by="lift"), 20)

inspect(top20rules.milk)

```

Graphical representation of rules

```{r}
plot(rules, cex=1)
plot(rules, measure=c("support","lift"), shading="confidence")

plot(top10rules, cex=1)

plot(top10rules, cex=1)

plot(top10rules, cex=1, method="matrix", measure = "lift")

plot(top10rules, method="grouped")

plot(top10rules, method="grouped", control=list(k=10), interactive = FALSE)

plot(top10rules, method="paracoord")

plot(top10rules, method="paracoord", control=list(reorder=FALSE))
 
plot(top10rules,method="graph", shading="lift")

```
