---
title: "Introduction to package *ggplot2* part 1"
author: "Mariusz Rafało"
output:
  html_document: default
  html_notebook: default
editor_options: 
  chunk_output_type: console
---

*Support material for class: Data Visualisation and Presentation at Warsaw School of Economics*

&copy; *Copyright Mariusz Rafało*

---

####Summary
This document presents basic concepts of decision tree analysis performed in *R*.

####Data
Dataset *landdata-states.csv* was downloaded from [Research consulting for the Harvard community](http://tutorials.iq.harvard.edu/R/Rgraphics/dataSets).

####Introduction 

Environment setup

```{r setup, include=FALSE, ECHO=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, echo=TRUE, warning=FALSE, message=FALSE)
```

Clear workspace
```{r}
rm(list=ls())
```

Libraries used
```{r}
library(ggplot2)
library(data.table) #fread
library(reshape2) #melt
library(magrittr)
library(dplyr)

```

Load data from *landdata-states.csv* file

```{r}
dane <- read.csv("E:/Data Sets/housing/landdata-states.csv", header=TRUE, sep=",")
d <- dane
```

####Basic analysis

Single variable (continuous)

```{r}
wyk.hist <- ggplot(dane, aes(x = Home.Value)) + geom_histogram(color="red", fill="green") 
wyk.hist

wyk.hist <- ggplot(dane, aes(x = Home.Value)) + geom_histogram(color="red", fill="red", binwidth = 5000) 
wyk.hist

wyk.hist + xlab("Value") + ylab("Quantity")

wyk.hist + xlab("Value") + ylab("Quantity") + ggtitle("Home price")

wyk.hist + xlim(c(0,500000))

qplot(data=dane, x=Home.Value, geom = "histogram")+ xlab("Value")

wyk.hist <- ggplot(dane, aes(x = Home.Value)) + geom_histogram(color="red", fill="red") 
wyk.hist + geom_freqpoly(color="blue")
wyk.hist
```

Two variables (continuous)

```{r}
dane.sub <- dane[dane$State %in% c("IL","LA","SD","CO"),]

wyk.sub <- ggplot(dane.sub,aes(x=Date,y=Home.Value,color=State))+geom_point()
wyk.sub

ggplot(dane.sub,aes(x=Date,y=Home.Value,color=State))+geom_point() + scale_y_log10()

ggplot(dane.sub,aes(x=Date,y=Home.Value,color=State))+geom_smooth()

ggplot(dane.sub,aes(x=Date,y=Home.Value,color=State))+geom_point()+geom_smooth()

```

Several variables on single plot

```{r}
wyk.bazowy <- ggplot(dane.sub, aes(x = Date , y = Land.Value/1000)) + xlab("Rok") + ylab("Land prices (kUSD)")

wyk.bazowy + geom_text(aes(label=State), size = 3)

wyk.bazowy + geom_point(aes(color = Home.Value)) 

wyk.bazowy + geom_point(aes(color=Home.Value/1000, shape = region), size=2) + scale_color_gradient(name="leg", low = "red", high = "yellow")


wyk.bazowy + geom_line(aes(color = State))


```

One continuous variable and several discrete variables

```{r}
ggplot(dane.sub, aes(x=State, y=Home.Value/1000)) + geom_bar(stat="identity") + ggtitle("Home prices")

ggplot(dane.sub, aes(x=State, y=Home.Value/1000)) + geom_boxplot()

ggplot(dane.sub, aes(x=State, y=Home.Value/1000)) + geom_violin(scale="area", color=12, alpha=0.5) +geom_jitter(alpha=0.5, aes(color=region),position = position_jitter(width = 0.1))

```

####Advanced analysis

3 variables (jitter)

```{r}

dane.sub <- dane[dane$State %in% c("IL","LA","SD","CO","SO","VA","WY"),]

wyk.jitter <- ggplot(dane.sub, aes(x = State,y = Home.Price.Index)) + 
        theme(legend.position="top",axis.text=element_text(size = 6))

wyk.jitter + geom_point(aes(color = Date), alpha = 0.4, size = 1.5,
                       position = position_jitter(width = 0.1, height = 0))

```


Heatmap

```{r}

ggplot(data = melt(cor(dane[,c("Home.Value","Land.Value","Land.Share..Pct.","Year")])), aes(x=Var1, y=Var2)) + geom_tile(aes(fill=value)) + xlab("") + ylab("")


ggplot(data = dane, aes(x=Home.Value, y=Land.Value)) + geom_hex() +
  scale_fill_gradient(low = "green", high = "red")


ggplot(data = dane, aes(x=Home.Value, y=Land.Value)) + geom_density2d()

ggplot(data = dane.sub, aes(x=Home.Value, y=Land.Value)) + geom_density2d(aes(color=region))

```

Gradient plots

```{r}
ggplot(data = dane, aes(x=Home.Value, y=Land.Price.Index)) + stat_density2d(aes(fill = ..density..), geom = "raster", contour = FALSE)


ggplot(data = dane, aes(x=Land.Price.Index, y=Home.Price.Index)) + stat_density2d(aes(fill = ..density..), geom = "raster", contour = FALSE) + xlim(0,3) + ylim(0,2) +
  scale_fill_gradient(low = "grey", high = "red") 


ggplot(data = dane, aes(x=Land.Price.Index, y=Home.Price.Index)) + geom_point(aes(alpha = Year)) + scale_alpha(range = c(0.1, 0.8))

ggplot(data = dane, aes(x=Land.Price.Index, y=Home.Price.Index)) + geom_point(aes(color = Year)) +   scale_color_gradient(low = "green", high = "red") 


ggplot(data = dane, aes(x=Land.Price.Index, y=Home.Price.Index)) + geom_point(aes(color = Year)) +   scale_color_gradient(low = "green", high = "red") + xlim(0,3) + ylim(0,2)

ggplot(data = dane, aes(x=Land.Price.Index, y=Home.Price.Index)) + geom_point(aes(color = Year)) +   scale_color_gradient(low = "green", high = "red") + xlim(0,1) + ylim(0.5,1.5)

ggplot(data = dane, aes(x=Land.Price.Index, y=Home.Price.Index)) + geom_point(aes(color = Year)) +   scale_color_gradient(low = "green", high = "red") + xlim(0,1) + ylim(0.5,1.5) + facet_wrap(~region, nrow=1)

```


####Sample report 1

```{r r1}
dd <- dane %>% 
  group_by(State, region) %>%
  summarise(srednia = mean(Home.Value)) 

pos <- dd[order(dd$region),]$State

ggplot(dd[order(dd$region),], aes(x=State, y=srednia/1000)) + 
  geom_bar(stat="identity", aes(fill=region)) +
  scale_x_discrete(limits=pos) + xlab("Year") + ylab("Average home price") +
  scale_fill_discrete(name="Region")

```

####Sample report 2

```{r r2}

ggplot(d) + geom_line(aes(x=Date, y=Home.Value/1000, color=State, linetype=region), lwd=1) +
  scale_linetype_discrete(name='Region') +
  scale_color_discrete(name='State') + xlab("Year") + ylab("Home price (1000 USD)") 
```

####Sample report 3

```{r r3}
ggplot(d[d$State %in% c("CT","AR"),]) + geom_point(aes(x=Date, y=Home.Value/1000, color=Land.Value/1000, shape=State),size=4) +
  scale_color_gradient(name='Land price', low='green', high='red') +
  scale_shape(solid=TRUE, name="State")+ xlab("Year") + ylab("Home prices (1000 USD)") 

```

####Sample report 4

```{r r4}

ggplot(dane[dane$region %in% c("N. East","West"),],aes(x=State, y=Home.Value/1000)) + geom_violin(scale="area") +
  geom_jitter(aes(color=region), alpha=0.4, position = position_jitter(width=0.1)) +
  facet_wrap(~region)+ xlab("State") + ylab("Home prices (1000 USD)") +
  scale_color_discrete(name="Region")
```

####Sample report 5

```{r r5}
ggplot(d, aes(x=State, y=Date)) + 
  theme(legend.position = "top") +
  geom_jitter(aes(color=Home.Value), position = position_jitter(width=0.2), alpha=0.6, size=14) +
  scale_color_gradient(low="green", high="red")

```

####Sample report 6

```{r r6}

d.mean <- dane[dane$region %in% c("N. East","West"),] %>%
  group_by(region, Date) %>%
  summarise(Home.Value=mean(Home.Value)/1000, Land.Value=mean(Land.Value)/1000)


ggplot(d.mean) + geom_hex(aes(x=Home.Value, y=Land.Value)) + facet_wrap(~region) + 
  scale_fill_gradient(low="green", high="red", name="Liczba") +
  xlab("Land price (1000 USD)") + ylab("Home price (1000 USD)")

```

####Sample report 7

```{r r7}

ggplot(dane, aes(x=Year, y=State)) + geom_tile(aes(fill=Home.Value/1000)) +
  scale_fill_gradient2(name="Home price (1000 USD)",low="yellow", mid="green", high="red", midpoint=median(d$Home.Value)/1000) +
  xlab("Year") + ylab("State")

```

####Sample report 8

```{r r8}

ggplot(d[d$State %in% c("CT","AR"),]) + 
  geom_point(aes(x=Date, y=Home.Value/1000, color=State, size=Land.Price.Index+5), alpha=0.4) +
   xlab("Year") + ylab("Home price (1000 USD)") +
  scale_color_discrete(name="State") + scale_size_continuous(name="Land price index") +
  theme(legend.position="bottom",axis.text=element_text(size = 16))

```
